---
title: "Phillips-curve model estimation"
#author: "Jørn I. Halvorsen"
output:
  html_notebook:
#    css: OFF.css
#    css: ON.css
    toc: true
    toc_depth: 4
---

```{r, eval=F, echo=F}
rmarkdown::render("Application_Phillips_static.Rmd")
```

# Data sample (realized DGP)
```{r message=FALSE, warning=FALSE}
#knitr::purl("Applications_Phillips_static.Rmd")
# Loading data
rm(list=ls())
library(sandwich)
library(lmtest)
library(wooldridge)
library(ggplot2)
library(plotly)
phillips_rdgp <- phillips # Realized DGP
#write.csv(phillips_rdgp, file='/home.csv')
```

From `r phillips_rdgp$year[1]` to `r rev(phillips_rdgp$year)[1]`

# Data generating process (DGP) and its regression model

Model specification 1:
\begin{equation}
inf_{t} = \beta_{0} +\beta_{1}unem_{t} + u_{t}, t=1,2,...,n
\end{equation}

```{r}
model_ph <- lm(phillips_rdgp$inf ~ phillips_rdgp$unem)
```

Model specification 2:
\begin{equation}
\Delta inf_{t} = \beta_{0} +\beta_{1} unem_{t} + u_{t}, t=1,2,...,n
\end{equation}

```{r}
model_phc <- lm(phillips_rdgp$cinf ~ phillips_rdgp$unem)
```

# Model estimation

```{r, echo=T, include=T}
ols_ph <- summary(model_ph)
ols_phc <-summary(model_phc)
```

# Diagnostics
## Detecting
### Time trends and/or weakly dependency (I(0) or I(1))

#### Graphical

```{r}
ggplot2::ggplot(data=phillips_rdgp) + ggplot2::geom_line(aes(x=year,y=inf))
ggplot2::ggplot(data=phillips_rdgp) + ggplot2::geom_line(aes(x=year,y=unem))
ggplot2::ggplot(data=phillips_rdgp) + ggplot2::geom_line(aes(x=year,y=cinf))
```

#### Formally

```{r eval=FALSE, include=FALSE}
phillips_rdgp$linf <- dplyr::lag(phillips_rdgp$inf)
phillips_rdgp$lunem <- dplyr::lag(phillips_rdgp$unem)
lm(phillips_rdgp$inf ~ phillips_rdgp$linf)
lm(phillips_rdgp$unem ~ phillips_rdgp$lunem)
```


### Serial correlation

$H_{0}:\rho=0$
$H_{1}:\rho\neq0$

```{r}
par(mfrow=c(1,2))
plot(model_ph$res)
plot(model_phc$res)
```

#### Option 1: t-test AR(1) on model 1 (with strictly exogenous regressors)

```{r}
## Serial correlation with strictly exogenous regressors
#### Step 1: Find the estimated residual
resval <- model_ph$res
#### Step 2: Estimat the estimated residual on itself
model_ar <- lm(resval~dplyr::lag(resval))
#### Step 3: Show the results from the estimation
summary(model_ar)
```

#### Option 2: Testing for serial correlation with general regressors on model 1 (with weakly exogenous regressors)

```{r}
## Perform Breusch-Godfrey test for first-order serial correlation:
bgtest(phillips_rdgp$inf ~ phillips_rdgp$unem)
```

# Estimation and output results for model 2

## Non-robust
```{r}
summary(model_phc, robust=FALSE)
```

## Robust [(HAC)](https://www.r-econometrics.com/methods/hcrobusterrors/) t-testing

```{r}
# Robust t test
coeftest(model_phc, vcov = vcovHC(model_phc, type = "HC0"))
```
