---
title: "Part 2 and 3: Exercises"
output:
  html_notebook:
    toc: true
    toc_depth: 3
---

- [Lenke til PDF-versjon løsningsforslag](https://github.com/joernih/MSB104TimeSeries/blob/main/inst/solutions/Exercises_solutions.pdf)

```{R eval=F, include=F}
rmarkdown::render("exercises.Rmd",output_format=rmarkdown::html_notebook(css="css/OFF.css"),output_file="exercises.Rmd")
rmarkdown::render("exercises.Rmd",output_format=rmarkdown::html_notebook(css="css/ON.css"),output_file="exercises.Rmd")
system('brave exercises.html')
```

```{r, eval=T, include=F}
#library(MSB104EconometricsJIH)
#data(package='MSB104EconometricsJIH')
```

# 10: Basic regression analysis with time series data 
1. Explain the nature of time series data

2. We have the following finite distributed lagged model:
\begin{equation}
y_{t}=\alpha+\delta_{0}z_{t}+\delta_{1}z_{t-1}+\delta_{2}z_{t-2}+u_{t}
\end{equation}
3. Show that a) Temporary change by 1 in period t can be written as:
- t periods: $\delta_{0}$
- t+1. periods: $\delta_{1}$
- t+2. periods: $\delta_{2}$
- t+3. periods: $0$

b) Permanent change by 1 in period t
- t periods: $\delta_{0}$
- t+1. periods: $\delta_{0}+\delta_{1}$
- t+2. periods: $\delta_{0}+\delta_{1}+\delta_{2}$
- t+3. periods: $\delta_{0}+\delta_{1}+\delta_{2}$

4. Why do think the estimation of such coefficients could be of interest to policy makers?
5. What is meant by that OLS is BLUE?
6. Is OLS of any use if (a) TS.3 is violated or (b) TS.5?
7. What is the problem with spurious regression?
8. How can the problem of spurious regression be detected and solved?
9. How would you interpret a dummy variables?
10. Why do seasonality occur in time series data?

# 11: Further issues in using OLS with time series data 

We have that:

- (a) MA(1) process: $y_{t}=\alpha_{}e_{t-1}+e_{t}$\
- (b) AR(1) process: $y_{t}=\rho y_{t-1}+e_{t}$, where we assume that $|\rho|<1$
- (c) Random walk: $y_{t}=\rho y_{t-1}+e_{t}$ \ 
- (d) Random walk with drift: $y_{t}=\alpha_{0}+\rho y_{t-1}+e_{t}$
Where $e_{t} \sim i.i.d. N(0,\sigma^2_{e})$

1. For (a), (b) find its (i) expected value and (ii) one period ahead forecast.
For (c) and (d) find its (i) expected value, (ii) variance, (iii) h-periods ahead forecast and (iv) correlation coefficient

2. What is the difference between weakly and strongly exogenous regressors? 

3. What is meant that time series is (a) $I(0)$ and (b) $I(1)$?

4. List some examples of time series that are (a) weakly dependent (b) highly persistent time series?

5. Why is the assumption of $u_{t} \sim  N(0,\sigma^2_{u})$ not needed in large sample for the  estimated coefficient to be (asymptotically) normally distributed?

# 12: Serial correlation and heteroskedasticity in time series regression 
1. What is the impact of serial correlation in a time series regression?
2. Is $R^{2}$ measured correctly under the presence of serial correlation?
3. What is meant by that OLS consistent but ~~not~~ biased?
4. Explain the difference between TS.1 and TS'1.
5. Why use robust serial correlation (HAC)?

# 16 Simultanous equations models (SEM) 
<br>
**Only for MSB104**

1. Explain the three different forms of endogeneity problem?
2. What is meant by a structural equation?
3. What is the problem of applying OLS to structural equation?
4. Which other method should be used and what is the identification condition for this to work properly (2-variable system)? 

```{r include=FALSE}
knitr::knit_exit()
```


