---
#title: "Application: Effects of personal exemption on fertility rates"
#author: "Jørn I. Halvorsen"
# output:
#   bookdown::html_document2:
#output:
#  html_document:
#    toc: false
#    toc_depth: 6
---

### 1. Data sample (realized DGP)
```{r message=FALSE, warning=FALSE}
# Loading data
rm(list=ls())
library(sandwich)
library(lmtest)
library(wooldridge)
library(ggplot2)
library(plotly)
fertility_rdgp <- fertil3 # Realized DGP
```

From `r fertility_rdgp$year[1]` to `r rev(fertility_rdgp$year)[1]`

### 2. Data generating process (DGP) and its regression model

Model specification 1:
\begin{equation}
gfr_{t} = \alpha_{0} + \delta_{0}pe_{t} + \delta_{1}pe_{t-1} + \delta_{2}pe_{t-1} + u_{t}, t=1,2,...,n
\end{equation}

```{R}
model_ferd <- lm(fertility_rdgp$gfr ~ fertility_rdgp$pe+ fertility_rdgp$pe_1+fertility_rdgp$pe_2)
```

Model specification 2:
\begin{equation}
\Delta gfr_{t} = \alpha_{0} + \delta_{0}\Delta pe_{t} + \delta_{1}\Delta pe_{t-1} + \delta_{2}\Delta pe_{t-1} + u_{t}, t=1,2,...,n
\end{equation}

```{R}
model_ferdc <- lm(fertility_rdgp$cgfr ~ fertility_rdgp$cpe +fertility_rdgp$cpe_1+fertility_rdgp$cpe_2)
```

### 3. Model estimation

```{r}
ols_ferd <- summary(model_ferd)
ols_ferdc <- summary(model_ferdc)
```

### 4. Diagnostics

#### Detecting

##### Time trends and/or weakly dependency (I(0) or I(1))

###### Graphical

Model specification 1:

```{r}
p1 <- ggplot2::ggplot(data=fertility_rdgp) + ggplot2::geom_line(aes(x=year,y=gfr))
p2 <- ggplot2::ggplot(data=fertility_rdgp) + ggplot2::geom_line(aes(x=year,y=pe))
p3 <- ggplot2::ggplot(data=fertility_rdgp) + ggplot2::geom_line(aes(x=year,y=pe_1))
p4 <- ggplot2::ggplot(data=fertility_rdgp) + ggplot2::geom_line(aes(x=year,y=pe_2))
gridExtra::grid.arrange(p1, p2, p3, p4,nrow = 1)
```
Model specification 2:

```{r}
cp1 <- ggplot2::ggplot(data=fertility_rdgp) + ggplot2::geom_line(aes(x=year,y=cgfr))
cp2 <- ggplot2::ggplot(data=fertility_rdgp) + ggplot2::geom_line(aes(x=year,y=cpe))
cp3 <- ggplot2::ggplot(data=fertility_rdgp) + ggplot2::geom_line(aes(x=year,y=cpe_1))
cp4 <- ggplot2::ggplot(data=fertility_rdgp) + ggplot2::geom_line(aes(x=year,y=cpe_2))
gridExtra::grid.arrange(cp1, cp2, cp3, cp4,nrow = 1)
```


###### Formally

```{r eval=FALSE, include=FALSE}
fertility_rdgp$lpe_1 <- dplyr::lag(fertility_rdgp$pe_1)
lm(fertility_rdgp$pe_1 ~ fertility_rdgp$lpe_1)
```


##### Serial correlation

```{r}
par(mfrow=c(1,2))
plot(model_ferd$res)
plot(model_ferdc$res)
```


Option 1: t-test AR(1) on model 1

```{r}
## Serial correlation with strictly exogenous regressors
#### Step 1: Find the estimated residual
resval <- model_ferd$res
#### Step 2: Estimat the estimated residual on itself
model_ar <- lm(resval~lag(resval))
#### Step 3: Show the results from the estimation
summary(model_ar)
```

Option 2: Testing for serial correlation with general regressors on model 1

```{r}
## Perform Breusch-Godfrey test for first-order serial correlation:
bgtest(fertility_rdgp$gfr ~ fertility_rdgp$pe_1+fertility_rdgp$pe_2)
```

### 7. Estimation and output results 
#### Non-robust
```{r}
summary(model_ferdc, robust=FALSE)
```

#### Robust t-testing (HAC)
```{r}
# Robust t test
coeftest(model_ferdc, vcov = vcovHC(model_ferdc, type = "HC0"))
```


